\begin{acknowledgements}

I am very grateful for all the contributions of co-authors and discussions with all of my colleagues over the last years.
I have learned a lot from you. While a thesis is traditionally the work of a single author, you are always included
whenever I used the term \emph{we}. I especially value the support and the very fruitful discussions with my
supervisors Tim Palmer and Peter D\"{u}ben, and appreciate the freedom to develop my own ideas.
For a self-propelled student it is easier to fly without headwind.

Particularly, I would like to thank 
Adam Paxton for countless efforts to correct my sloppy approach to mathematics;
Mosè Giordano for teaching me many tricks about Julia;
Matteo Croci for many discussions how to make stochastic rounding faster and faster;
Debbie Hopkins for having improved my general style of writing when we wrote our comment together;
Valentin Churavy and Jeffrey Sarnoff for Julia-related bug fixes;
Simon MacIntosh-Smith for access to the A64FX processors on Isambard 2;
Sam Hatfield and Josh Dorrington for RaspberryPi and Google challenges in my favourite office;
Simon Proud and Lucy Harrington for providing further data sets to test our compression method;
Miha Razinger and Juanjo Dominguez for being great mentors during two ESoWCs;
Peter Coveney for writing countless emails to discuss periodic orbits;
Chiara Maiocchi for explaining unstable periodic orbits to me;
Joe Bates for having introduced me to logarithmic fixed-point numbers;
Myles Allen and David Lee for supervising me on decarbonising aviation; 
Sören Thomsen for many initiatives to decarbonise science together;
Mike Morrison for great ideas on how to make science user-friendly;
and
Najar's for having fed me countless falafels over the last years whenever my work rhythm was very offbeat.

Special thanks to the editors and many anonymous reviewers that have read and improved our manuscripts over the
last years without getting appropriate credit for these efforts --- particularly also the reviewers that took the time to
formulate appreciation among all the constructive criticism. Thanks for your contribution to make science more 
welcoming to students and early career scientists that do not always submit perfect manuscripts. Many thanks also
to Geoff, Lukas, Chloë and Žiga for proofreading my thesis.

I would like to thank the whole Julia community for an uncountable effort to develop a programming language that 
turned out to be the secret behind much of my progress over the last years. I have benefited a lot from learning
Julia and I hope I was able to return the favour through the development and contribution to several Julia packages.
Many thanks also to everyone who developed the matplotlib plotting library, which was used for every figure in this thesis,
and taught me a lot about data visualisation.

Over the last four years I have received funding from various sources which I am very grateful for. The European
Research Council and therefore indirectly the European Union for my stipend; the Natural Environmental Research
Council for my tuition fees, the Copernicus Programme and the ECMWF Summer of Weather Code 2020 and 2021
and therefore indirectly the European Commission for funding my research on climate data compression; Jesus
College for several travel grants that supported my attendance at conferences; and a big greedy publisher for the first
prize at a student conference --- thanks for the money, but I still will not publish with you.

Then I am very grateful to all the more than 30 people I had the opportunity to live with in the Vic over the last four years.
You've always been like a family to me; I've tried to invest a lot of time and energy into this amazing community and you
have always more than returned the favour with laughter and calories, with support and activity, with challenge and
knowledge, with lessons and help, with Gwen and Winston, and definitely turned me into a better human.

Finally, I would like to thank my mum for always having worked hard to support me towards the opportunities to do what
I want to do.
 
\end{acknowledgements}
